Embedded Scalars, Preferred Readings \& Prosody: {A}n Experimental Revisit

### 2015-02-04 text/paper/revision_01/paper.tex
K2 revision reflection

To check whether the model yields sensible results we first look at
the preference-related control conditions.
%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{pics/post_salience_TF.pdf}
  \caption{Estimated posteriors for the salience of \lc- and
    \ec-readings for different prosodic cues. Lines under curves
    indicate 95\% HDIs (see Footnote~\ref{fn:HDI}).}
  \label{fig:Posterior_TF}
\end{figure}
%
Estimates for the (marginalized) posteriors over relative salience
levels for target-related controls are plotted in
Figure~\ref{fig:Posterior_TF}. The results are exactly as we would
expect them to be. Given our data, we should believe that the
\lc-reading is most prominent. The level of prominence varies with
accentuation. Under prosody that we hypothesized would favor the
\lc-reading, the contrast is most pronounced. Under accentuation that
we hypothesized favors the \ec-reading we are no longer justified in
believing with certainty that the \lc-reading is preferred (by
comparison of 95\% HDIs, see below).\footnote{\label{fn:HDI} A 95\%
  Highest Density Interval (HDI) is a convex region of values over
  which 95\% of the distribution's probability density is distributed,
  such that no value outside of that region has higher probability
  density than any point within
  \citep{Kruschke2011:Doing-Bayesian-}. Intuitively speaking, the 95\%
  HDI is the set of values which we may believe to be true with some
  confidence; what falls outside this region is what can be excluded
  safely.}

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{pics/post_salience.pdf}
  \caption{Posteriors over relative salience of target readings in
    critical conditions.}
  \label{fig:Posterior_T}
\end{figure}

@@

\paragraph{Summary \& Interpretation.} The present study investigated
prosodic effects on quantifier interpretation by employing a new
experimental paradigm, the incremental verification
task. Based on a current debate in the literature, we were
specifically interested in whether ``local readings'' of sentences
where \emph{some} is embedded under another quantifier can be found,
and how salient these readings are under varying prosody of auditorily
presented sentence material, in particular the presence or absence of
contrastive stress on embedded \emph{some}.

@@

In sum, the present findings cannot be explained in terms of the
austere and stylized versions of traditionalism and grammaticalism
that we have set up so far. But we should ask what plausible
additional assumptions or alternative auxiliary assumptions would
enable either core theory to accommodate the data \emph{ex post}. This
is what we do in the next section, where we also reflect critically on
our design and our interpretation of the data.

\section{Critical reflection}
\label{sec:critical-reflection}

\paragraph{Alternative explanations based on traditionalism and
  grammaticalism.} Section~\ref{sec:theories-predictions} introduced
traditionalism and grammaticalism as two competing core theories,
whose ability to make empirical prediction depends on additional
auxiliary assumptions. We choose, in part for practical reasons,
particular auxiliary assumptions, but we should also ask more
generally: on the supposition that our design and the offered
interpretation of our data are sound, what would it take to
accommodate the observed data under the core theories that we
considered?

Consider grammaticalism first. Our data contradicts the notion that
preferences follow logical strength. But, as noted in
Section~\ref{sec:theories-predictions}, grammaticalism could be
supplemented by a conceptually different disambiguation
criterion. Selecting readings in terms of how well they answer the
contextually given question under discussion, as suggested by
\citet{Fox2007:Free-Choice-and,GualminiHulsey2008:The-Question-An}, is
an option. But it is quite unclear what the question under discussion
should be that guided judgements in our particular task. Similarly, it
is not impossible that \citeauthor{Magri2011:Another-Argumen}'s
(\citeyear{Magri2011:Another-Argumen}) proposal involving obligatory
exhaustification and variable alternatives that feed into it can
accommodate the observed data. \todo{look at reviewer comment and
  acknowledge duly.}

However, looking at the qualitative pattern in
(\ref{bsp:observed-preferences}), a \emph{prima facie} plausible
disambiguation criterion is ready at hand: if we assume that each
insertion of the $\exh(\cdot)$ operator is \emph{costly}
anti-proportional to the depth of its embedding under other logical
operators, then we derive a preference of literal readings over local
ones over global ones. Whether such an \emph{economy principle} is
sound and in line with further empirical data would then be an
interesting matter of further investigation. What this economy
principle does not explain yet is the quantitative differences in the
preference orders between \as- and \es-sentences. Still, if combined
with further information about the global properties of the sentences
in question, this might be an interesting alternative disambiguation
criterion for an empirically testable grammaticalism.

Traditionalism, on the other hand, has to explain the possibility of
local readings of prosodically unmarked \es-sentences, and, perhaps,
also the large number of local readings in accented \es-sentences
compared to the much lower number in accented \as-sentences. To do so
within a strict traditionalist framework seems difficult. A first
option might be to account for alleged local readings as the result of
some other phenomenon, such as typicality or contrast
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar,Tielvan-Tiel2014:Quantity-Matter}.
Since this ties into a critical reflection on our design, we will
enlarge on the possibility of typicality- and contrast-based effects
in our setting. Eventually, we argue that contrast-effects alone are
unlikely to explain the observed high number of local answers in
\es-sentences, but that a ``modern traditionalist'' explanation of the
answer pattern might be available if we allow for the possibility that
\emph{exactly one} gets an unexpected ``referential interpretation.''

@@

To wrap up, we suggested amendements to grammaticalism and
traditionalism that could help either position explain the data we
observed. Whether these suggestions can be vindicated by further
empirical research is a matter that we must leave open. It remains,
however, that our data provides interesting challenges to both
traditionalism and grammaticalism.

% \paragraph{Critique.} There are a number of conceivable objections to
% our design and our interpretation of the presented results. We will
% consider the most pressing ones in the following.

>>>>

To check whether the model yields sensible results we first look at
the preference-related control conditions.
%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{pics/post_salience_TF_small.png}
  \caption{Estimated posteriors for the salience of \lc- and
    \ec-readings for different prosodic cues. Lines under curves
    indicate 95\% HDIs (see Footnote~\ref{fn:HDI}).}
  \label{fig:Posterior_TF}
\end{figure}
%
Estimates for the (marginalized) posteriors over relative salience
levels for target-related controls are plotted in
Figure~\ref{fig:Posterior_TF}. The results are exactly as we would
expect them to be. Given our data, we should believe that the
\lc-reading is most prominent. The level of prominence varies with
accentuation. Under prosody that we hypothesized would favor the
\lc-reading, the contrast is most pronounced. Under accentuation that
we hypothesized favors the \ec-reading we are no longer justified in
believing with certainty that the \lc-reading is preferred (by
comparison of 95\% HDIs, see below).\footnote{\label{fn:HDI} A 95\%
  Highest Density Interval (HDI) is a convex region of values over
  which 95\% of the distribution's probability density is distributed,
  such that no value outside of that region has higher probability
  density than any point within
  \citep{Kruschke2011:Doing-Bayesian-}. Intuitively speaking, the 95\%
  HDI is the set of values which we may believe to be true with some
  confidence; what falls outside this region is what can be excluded
  safely.}

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{pics/post_salience_small.png}
  \caption{Posteriors over relative salience of target readings in
    critical conditions.}
  \label{fig:Posterior_T}
\end{figure}

@@

\paragraph{Summary \& Interpretation.} The present study investigated
prosodic effects on quantifier interpretation by employing a new
experimental paradigm, the incremental verification task. We were
specifically interested in whether ``local readings'' of sentences
where \emph{some} is embedded under another quantifier can be found,
and how salient these readings are under varying prosody of auditorily
presented sentence material, in particular the presence or absence of
contrastive stress on embedded \emph{some}.

@@

In sum, the present findings cannot be explained in terms of the
versions of traditionalism and grammaticalism that we have considered
so far, i.e., with the particular auxiliary assumptions adopted in
Section~\ref{sec:theories-predictions}. But we should ask what
plausible amendements or alternative auxiliary assumptions would
enable either core theory to accommodate the data \emph{ex post}. This
is what we do in the next section, where we also reflect critically on
our design and our interpretation of the data.

\section{Reflection}
\label{sec:critical-reflection}

\subsection{Alternative explanations within the core theories}

Section~\ref{sec:theories-predictions} introduced
traditionalism and grammaticalism as two competing core theories,
whose concrete empirical predictions depend on additional auxiliary
assumptions. We chose, in part for practical reasons, particular
auxiliary assumptions, but we should also ask more generally: on the
supposition that our design and the offered interpretation of our data
are sound, what would it take to accommodate the observed data under
the core theories?

\paragraph{Grammaticalism.} Consider grammaticalism first. Our data
contradicts the notion that preferences follow logical strength. But,
as noted in Section~\ref{sec:theories-predictions}, grammaticalism
could be supplemented by a conceptually different disambiguation
criterion. Selecting readings in terms of how well they answer the
contextually given question under discussion, as suggested by
\citet{Fox2007:Free-Choice-and,GualminiHulsey2008:The-Question-An}, is
an option. But it is quite unclear what the question under discussion
should be that guided judgements in our particular task. To make this
approach yield \emph{ex ante} predictions, we would need to determine
before the data is available which implicit QUDs our particular task
should give rise to, so as to derive a preference ordering over
readings. This may well be possible, but we do not presently see how.

Another possibility of explaining our data within a grammaticalist
core theory is to adopt \citeauthor{Magri2011:Another-Argumen}'s
(\citeyear{Magri2011:Another-Argumen}) proposal that exhaustification
operators occur at every relevant scope site while alternative sets as
their input may also be empty.\footnote{This interesting line of
  alternative \emph{post hoc} explanation was first suggest by a
  reviewer, but worked out slightly differently here to strengthen the
  reviewer's case.} For \as- and \es-sentences, we would have to
consider the parses in (\ref{Magri-Parses}) (in simplified notation).

\begin{exe}
  \ex \label{Magri-Parses}
    \begin{xlist}
    \ex \label{Magri-Parse-AS} \mymark{$\exh_{Alt_1}$}(\mymark{All} $x$ are such that \mymark{$\exh_{Alt_2}$}( $x$ is connected to
      \mymark{some}  \dots.))
    \ex \label{Magri-Parse-ES} \mymark{$\exh_{Alt_1}$}(\mymark{Exactly
        one} $x$  is such that \mymark{$\exh_{Alt_2}$}( $x$ is connected to \mymark{some}  \dots.))
    \end{xlist}
\end{exe}

\noindent The alternative sets $Alt_{1,2}$ are plausibly either both
empty or both non-empty. In the former case, we get the literal
reading. In the latter case, we should assume that alternatives are
just the standard ones that asymmetrically entail the literal reading:

\begin{exe}
  \ex \label{Magri-Alternatives} 
    \begin{xlist}
    \ex \label{Magri-Alternatives-AS} For \as-sentences: \\
      $Alt_1 = \set{\text{``All } x
          \dots x \text{ is connected to all } \dots \text{''}}$ \\
        $Alt_2 = \set{\text{``}x \text{ is connected to all } \dots
          \text{''}}$
    \ex \label{Magri-Alternatives-AS} For \es-sentences: \\
      % $Alt_1 = \set{\text{``Exactly 1 } x
      %     \dots x \text{ is connected to all } \dots \text{''}}$ \\
      $Alt_1 = \emptyset$ \\
        $Alt_2 = \set{\text{``}x \text{ is connected to all } \dots \text{''}}$
    \end{xlist}
\end{exe}

\noindent Notice that $Alt_1$ is empty for \es-sentences if we
restrict attention to the ``global alternatives'' that asymmetrically
entail the to-be-interpreted sentence. Interestingly, the readings
that this approach derives are exactly the literal readings (under
empty alternatives) and the local readings (under non-empty
alternatives). What is left to explain is the preference relation over
these readings. Here, the Magri-style approach could resort to
considerations of \emph{parsing economy}: reasoning with fever
alternatives is easier than reasoning with more alternatives. This
would explain the preference for literal readings and also why local
readings are more strongly attested for \es-sentences.

On conceptual grounds we are very much in favor of a disambiguation
criterion in terms of economy considerations, but whether the
particular account sketched here is empirically successful in other
cases as well, must remain to be seen. Two things are worth
emphasizing here nonetheless. Firstly, the sketched Magri-style
account is a plausible alternative \emph{post hoc} explanation. It was
not, at least to our knowledge, a salient possibility before seeing
our data set: other specifications of $Alt_1$ and $Alt_2$ could have
been made equally plausible \emph{ex ante}. Secondly, what our data
refutes is that preferences for readings follow logical
strength. This, therefore, does not jeopardize grammaticalism as a
core theory, but calls for theory-internal revision of its
disambiguation criterion. Taken together, we argue that the data
presented here points to a dire need of refining grammaticalism,
perhaps along the lines suggested here, but ideally alongside more
empirical data like the one we presented.

\paragraph{Traditionalism.} Traditionalism, as we described it in
Section~\ref{sec:theories-predictions}, is also uncomfortable with our
data. It has to explain the possibility of local readings of
prosodically unmarked \es-sentences, and, perhaps, also the large
number of local readings in accented \es-sentences compared to the
much lower number in accented \as-sentences. To do so within a strict
traditionalist framework seems difficult. A first option might be to
account for alleged local readings as the result of some other
phenomenon, such as typicality or contrast
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar,Tielvan-Tiel2014:Quantity-Matter}.
Since this ties into a critical reflection on our design, we will
enlarge on the possibility of typicality- and contrast-based effects
in our setting. Eventually, we argue that contrast-effects alone are
unlikely to explain the observed high number of local answers in
\es-sentences, but that a ``modern traditionalist'' explanation of the
answer pattern might be available if we allow for the possibility that
\emph{exactly one} gets an unexpected ``referential interpretation.''

@@

To wrap up, we suggested amendements to grammaticalism and
traditionalism that could help either position explain the data we
observed. Whether these suggestions can be vindicated by further
empirical research is a matter that we must leave open. It remains,
however, that our data provides interesting challenges to both
traditionalism and grammaticalism, without necessarily refuting either
core theory as such.

\subsection{Critique}

The foregoing interpretation and discussion presupposed that the
interpretation of our data, given in Section~\ref{sec:discussion-1},
was sound. But there are a number of conceivable objections to our
design and our interpretation of the presented results. We will
consider the most pressing ones in the following.
